---
title: "CompositionalWords"
author: "Dhruv Mishra"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
setwd("/Users/dhruv/code/uni/assigments/rprogramming")
```

## Read the data
```{r}
comp.ratings <- readLines("data/comp-ratings.txt") %>%
  lapply(function(l) {parts = unlist(strsplit(l, "\t")); 
  paste(gsub("-n", "", parts[1]), gsub(" ", "\t", parts[2]), sep="\t")}) %>%
  unlist() %>% textConnection() %>% read.table(sep="\t", header=FALSE, row.names=NULL)

colnames(comp.ratings) <- c("Compound","Word1_mean","Word1_std","Word2_mean",
                            "Word2_std","Cpd_mean","Cpd_std","mean1*mean2")

comp.properties <- read.csv("data/comp-properties.txt", sep = "\t", 
                            row.names=NULL, na.strings=c("NA", "-", "?", "None"))

co.occ.compounts <- read.table("data/co-occ-compounds1.txt", header=FALSE, 
                               sep = "\t", stringsAsFactors=FALSE, row.names=NULL)

co.occ.constituents <- read.table("data/co-occ-constituents1.txt", header=FALSE,
                                  sep = "\t", stringsAsFactors=FALSE,
                                  row.names=NULL)

abstract.concrete <- read.csv("data/abstract-concrete.txt", sep = "\t")
```

## Descriptive Statistics and Plots
### Compositionality ratings
```{r}
summary(comp.ratings)

g <- ggplot(comp.ratings, aes(Word1_mean, Cpd_mean))
g + geom_jitter() + geom_smooth()

g <- ggplot(comp.ratings, aes(Word2_mean, Cpd_mean))
g + geom_jitter() + geom_smooth()
```

### Compositionality properties
```{r}
summary(comp.properties)
```


## The degree of compositionality of the English noun compunds is related to
### 1.1 The frequencies of the compounds
```{r}
comp <- inner_join(comp.ratings, comp.properties, by="Compound")
test <- t.test(comp$Cpd_mean, comp$FreqComp)
test
```
As can be seen, the confidence Interval of the diffence in mean of `Cpd_mean` and `FreqComp` is $`r test$conf.int`$. Since the p-value is very small (<0.05), we reject the null hypothesis. Therefore, **Compositionality is related to Compound frequency**.

### 1.2 The frequencies of the modifiers
```{r}
test <- t.test(comp$Cpd_mean, comp$FreqMod)
test
```
As can be seen, the confidence Interval of the diffence in mean of `Cpd_mean` and `FreqMod` is $`r test$conf.int`$. Since the p-value is very small (<0.05), we reject the null hypothesis. Therefore, **Compositionality is related to Modifier frequency**.

### 1.3 The frequencies of the modifiers
```{r}
test <- t.test(comp$Cpd_mean, comp$FreqHead)
test
```
As can be seen, the confidence Interval of the diffence in mean of `Cpd_mean` and `FreqHead` is $`r test$conf.int`$. Since the p-value is very small (<0.05), we reject the null hypothesis. Therefore, **Compositionality is related to Head frequency**.

**Therefore Compositionality is related to frequencies of compounds, modifiers and heads.**

### 2.1 The productivities of modifiers
```{r}
test <- t.test(comp$Cpd_mean, comp$ProdMod)
test
```
As can be seen, the confidence Interval of the diffence in mean of `Cpd_mean` and `ProdMod` is $`r test$conf.int`$. Since the p-value is very small (<0.05), we reject the null hypothesis. Therefore, **Compositionality is related to productivity of Modifier**.

### 2.2 The productivities of heads
```{r}
test <- t.test(comp$Cpd_mean, comp$ProdHead)
test
```
As can be seen, the confidence Interval of the diffence in mean of `Cpd_mean` and `ProdHead` is $`r test$conf.int`$. Since the p-value is very small (<0.05), we reject the null hypothesis. Therefore, **Compositionality is related to productivity of Head**.


### 3.1 The ambiguities of modifiers
```{r}
test <- t.test(comp$Cpd_mean, comp$AmbMod)
test
```
As can be seen, the confidence Interval of the diffence in mean of `Cpd_mean` and `AmbMod` is $`r test$conf.int`$. Since the p-value is very small (<0.05), we reject the null hypothesis. Therefore, **Compositionality is related to ambiguity of Modifier**.


### 3.2 The ambiguities of heads
```{r}
test <- t.test(comp$Cpd_mean, comp$AmbHead)
test
```
As can be seen, the confidence Interval of the diffence in mean of `Cpd_mean` and `AmbHead` is $`r test$conf.int`$. Since the p-value is very small (<0.05), we reject the null hypothesis. Therefore, **Compositionality is related to ambiguity of Heads**.

## The degree of compositionality of the English compound ratings is related to their degree of concreteness/abstractness.
```{r}
ac <- inner_join(comp.ratings, abstract.concrete, by=c("Compound" = "Word"))
test <- t.test(ac$Cpd_mean, ac$Conc.M)
test
```
As can be seen, the confidence Interval of the diffence in mean of `Cpd_mean` and `Conc.M` is $`r test$conf.int`$. Since the p-value is very small (<0.05), we reject the null hypothesis. Therefore, **Compositionality is related to the corresponding degree of concreteness/abstractness**.
